Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-06-17 Ben_Manuscripts/structure_paper/Draft.tex
Incorporated comments

  Nanostructured porous membranes made from the cross-linked hexagonal phase of
  self-assembled lyotropic liquid crystals (LLCs) are a promising material for
  selective separations. In this work, we investigate an experimentally characterized 
  LLC membrane using an atomistic molecular model. 
  %MRS8: I think this is a little too subtle a point to be making for the abstract. 
  %I added a bit different spin on this at the end of the abstract to gesture towards the idea. Thoughts?
  %While it is relatively 
  %straightforward to create a stable hexagonal phase using atomistic modeling, we demonstrate that 
  %closely reproducing the experimental pore structure requires a carefully 
  %constructed initial configuration in order to avoid falling into metastable
  %basins. 
  We use simulated X-ray diffraction (XRD) patterns in order to ensure 
  maximum consistency between our structures and experiment. We show that pores 
  are likely composed of 5 columns of stacked LLC monomers which surround each
  hydrophilic core. Although this system has been reported as dry, we show that
  small amounts of water are necessary to fully reproduce the experimental XRD
  pattern due to asymmetries introduced by hydrogen bonds between monomer head
  groups and water molecules. We explore the composition and structure of the 
  nanopores and reveal that there exists a composition gradient rather than a 
  abrupt partition between the hydrophilic and hydrophobic regions. A caveat is 
  that the time scales of the dynamics are extremely long for this system, 
  requiring starting from a range of alternative initial configurations and 
  careful examination of the metastable states observed.
  The clearer picture of the nanoscopic structure of these membranes provided in
  this study will enable a better understanding of the mechanisms of small 
  molecule transport within these nanopores.
  \end{abstract}

@@

% MRS6: I would keep a paragraph about here how current membranes DON'T
% work for this purpose. But one paragraph is enough.
% BJC5: added back below paragraph.  
% MRS7: Do a pass for clarity at the sentence level.  For example ``degree of stochasticity'' is vague.
% MRS8: ``because they are unstructured'' lacks detail.  NF has some structure, since there are irregular pores. Check some other reviews and see the points they cover?
  Current commercial RO and NF membranes suffer limitations to their selectivity
  because they are unstructured. Although scalable, their fabrication involves the 
  spontaneous assembly of polymers into disordered structures which offer separation 
  pathways that are tortuous and polydisperse in size. This makes overcoming the well-known 
  permeability-selectivity tradeoff a challenge. Namely, it is difficult to increase
  the permeability of a desired molecular or atomic species, while maintaining the 
  same retention of an undesired species \cite{werber_materials_2016}.

@@

%MRS8: preliminary evidence has shown they may be?  It's more than a guess.
  Crosslinked lyotropic liquid crystal (LLC) membranes may be capable of
  performing highly selective separations. LLCs are amphiphilic molecules that
  have the ability to self-assemble into porous nanostructures
  \cite{smith_ordered_1997} and can be crosslinked to create mechanically
  strong membrane films with pores on the order of 1 nm in diameter
  \cite{zhou_supported_2005}. Unlike most commercial NF membranes, LLC membrane pores
  are uniform in size because they are self-assembled. 
  Since LLC membranes lack a pore size distribution, they
  inherently exhibit high selectivity due to their strict molecular weight
  cut-off (MWCO)~\cite{zhou_supported_2005}. Additionally, the LLC monomers
  examined in this paper are salts, and therefore lead to Donnan exclusion of ions
  in solution. The membrane gains a net surface charge when counterions from the head
  groups that line the pore walls escape into the feed solution in an effort to balance 
  the gradients of concentration and electric potential \cite{donnan_theory_1995}.

@@

  Recently, researchers have learned how to macroscopically align the hexagonal
  domains which has revived research into H\textsubscript{II} phase LLC membranes.
  Previously, research efforts were focused on the Q\textsubscript{I} phase,
  whose geometry does not require alignment. In 2014, Feng et al.~showed that one
  can align Col\textsubscript{h} hexagonal domains, created by the monomer Na-GA3C11,
  using a magnetic field with subsequent cross-linking to lock the structure in 
  place\cite{feng_scalable_2014}. In 2016, Feng et al. showed that one could obtain
  the same result using a second technique termed soft confinement\cite{feng_thin_2016}.
  %MRS8: maybe say what it is, not just what it's named? 
  Current experimental efforts are focused on extending the method to the 
  H\textsubscript{II} phase and characterizing the performance of these newly aligned systems.

@@

  %BJC6: is this too bold considering the work done at University of Wisconsin?
  %MRS8: Maybe take out significantly
  In this study, we achieve a %significantly 
  more realistic atomistic description of LLC
  membranes than, to our knowledge, has ever previously been created, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the results using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure drawn from small angle X-ray scattering (SAXS)
  and wide angle X-ray scattering (WAXS) experiments as well as in matching ionic
  conductivity measurements \cite{feng_thin_2016}.

@@

  In this paper, we perform molecular modeling of the Col\textsubscript{h} assembly 
  formed by Na-GA3C11. Compared to the H\textsubscript{II} phase, the Col\textsubscript{h}
  phase is a simpler starting point. The system 
  %MRS8: check rephrasing on this, since we do say later it does have some water
  %has no water 
  is not assembled in aqueous phase
  which allowed us to simulate longer timescales, and there exists detailed experimental
  characterization of the fully aligned state, including 2D wide-angle X-ray
  scattering (WAXS) patterns (Figure~\ref{fig:WAXS}) which are useful for reconstructing 
  structural data. 
  %MRS8: I started tweaking the below, but it might be better just to eliminate.
  % Modeling of Col\textsubscript{h} will help in modeling 
  %H\textsubscript{II} in the future.
  %The two phases appear to share similar structural characteristics since the
  %pore spacings in each system are in close agreement
  %\cite{feng_thin_2016,resel_structural_2000}.

@@

  Despite having structural data, there is still information which experiment cannot
  definitively answer. 
  %MRS8: grammatically a bit redundant.
  %There are a number of structural questions we wish to answer in this article. 
  Specifically, we want to know:
  \begin{enumerate}
    \item What is the density of monomers that pack around each hydrophilic core? 
    \label{point:monomernum}

@@

    %MRS7: is this really a liquid crystalline system, even if it is made of liquid crystal monomers? Should we say
    % in many liquid crystal systems?
    %BJC5: changed to 'many liquid crystalline systems'
 	We are careful to avoid the term 'layers' since many liquid crystalline systems
 	have long range order in 1 or 2 spatial dimensions and short range order in 	
 	the other dimensions.~\cite{chaikin_principles_1995}  % page 58
 	In the system we are studying, there are long-range 2D correlations in the 
 	hexagonal array of pores (xy plane) and short range z-direction correlations between 
        stacked columns of monomers. In this study, we use atomistic molecular modeling
        to study how the system's structure is affected by the density of these monomer columns
        that surround each pore's hydrophilic core.

@@

	While the Col\textsubscript{h} phase is described as dry, it is likely that small amounts of
	ambient water are 
        %leached 
        absorbed
        into the system. The hydrogen bonding network formed by the 
	water may play a role in structuring the pore. We used simulated X-ray diffraction (XRD) patterns
	to see if there is any meaningful structural difference between a "dry" and a "wet" system.

	\item What is the 
          %MR8: chemical composition is a bit ambiguous; implies mole fractions, which isn't what we mean
          %chemical composition 
          detailed atomistic structure of the pores?\label{point:composition}

@@

  \subsection{Equilibration}
  %MRS7: could put this mostly in supporting. 
  %BJC6: cut in half
  We developed equilibration schemes for creating dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. For wet systems, we added the desired concentration of water to 
  the initial configuration and carried out equilibration in the same way
  as the dry systems. First, we fixed monomer head groups in place using position
  restraints with a force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We gradually
  released the position restraints by decreasing the force constants over a series
  of NVT simulations. We allowed the resulting unrestrained structure to equilibrate for
  5 ns in the NPT ensemble with pressure controlled by the berendsen barostat 
  followed by NPT equilibration simulations run for at least 400 ns using the 
  Parrinello-Rahman barostat. 
  %More precise details are 
  More equilibration details are 
  available in 
  section S-\ref{S-section:equilibration}.

@@

  % BJC5: trying things out with 'columns per pore' instead of 'monomers per layer'.
  Our simulations best support a model built with 5 monomer columns per pore based on
  the measured equilibrated pore-to-pore distances. To discern the composition of
  the monomer layers, addressing question \ref{point:monomernum} above, we ran
  simulations of systems created with 4--8 columns per pore. We built systems in both
  the parallel displaced and sandwiched configurations and equilibrated them according
  to the dry equilibration procedure. We tested all systems with an initial vertical monomer 
  spacing, $\mathit{d}$, of 3.7 \AA~in accordance with R-$\pi$. We tested 4 additional 
  systems with monomers initially spaced 5 \AA~apart vertically (see section~\ref{S-section:initial_pore_spacing}
  of the Supporting Information for more details on sensitivity to initial layer spacing). 
  We considered the pore-to-pore spacing to be equilibrated as defined in section~\ref{method:equil_time}.
  Figure~\ref{fig:p2p} shows the equilibrated pore-to-pore distances for all systems tested.

@@

  All systems tested, although equilibrated from the perspective of the metrics used here, are
  frozen in metastable basins. Not all make physical sense or fit the experimental profile that 
  we are trying to match. In the limit of infinite simulation time, all systems will in theory 
  converge to a single equilibrium configuration, but that time is far beyond the 100's of 
  nanosecond simulated here. For simplicity, we group the systems studied here into the ordered
  and disordered basins. What we find is that any system where $\mathit{d}=3.7$ \AA~can generally 
  be characterized as being in an
  more ordered basin, and any system where $\mathit{d}=5.0$ \AA~can be characterized as a more the disordered basin.
  Generally, when monomers are started further apart, they stay further apart than systems where
  monomers are started closer together (see Table~\ref{table:correlation_length}). Because of the
  extra space between stacked monomers, monomer head groups have more rotational freedom. We
  quantified the ordering of the head groups using the nematic order parameter (see section S~\ref{method:nematic_order} for details of calculation). 
  %MRS8: the reference above isn't appearing. I'm not sure why after quick check.
  Disordered basin systems have a lower nematic order parameter (meaning they are more disordered) 
  than ordered basin systems.

@@

  To understand the %true 
  more likely origin of R-spots, we determined which atoms gave rise to the 
  feature. 
  %MRS8: next two sentences repetitive; could you revise? 
  Since R-spots is present as higher intensity spots within R-alkanes, it is
  likely that the spots arise as a consequence of the tails. By removing all non-tail atoms
  from the trajectory and simulating a diffraction pattern with the remaining atoms, we were
  able to isolate the cause of the spots to the tails (Figure~\ref{fig:tails_rzplot}). Since the 
  tails stay nearly flat, we plotted their centroids (Figure \ref{fig:centroids})
  and measured the angle between each centroid and its nearest neighbors with respect to the 
  plane of the membrane. We see distinct peaks in the distribution of these angles 
  (Figure~\ref{fig:tail_packing}).

@@

%BJC6: figures are in order that they are talked about in text.
  \begin{figure}[!htb]
  \begin{subfigure}{0.32\linewidth}
  	\centering
  	\includegraphics[width=\textwidth]{rzplot_layered_280K_jet.png}
  	\caption{}\label{fig:sandwiched280K}
  \end{subfigure}
  \begin{subfigure}{0.32\linewidth}
    \includegraphics[width=\textwidth]{tilt_dist.png}
    \caption{}\label{fig:tilt}
  \end{subfigure}
  \begin{subfigure}{0.32\linewidth}
	\centering
	\includegraphics[width=\textwidth]{tails_rzplot_jet.png}
	\caption{}\label{fig:tails_rzplot}
  \end{subfigure}
  %BJC6: Brought this back to make 6 figures.
  \begin{subfigure}[t]{0.32\linewidth}
    \centering
	\includegraphics[width=\textwidth]{centroids_box.png}
  \caption{}\label{fig:centroids}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\linewidth}
        \centering
	        \includegraphics[width=\linewidth]{angles_traj_layered.png}
	        \caption{}~\label{fig:layered_tails}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
        	\centering
	        \includegraphics[width=\linewidth]{layered_angle_v_I.png}
	        \caption{}~\label{fig:layered_integration}
  \end{subfigure}
  %MRS8: nicely organized figure! I think in (c), you do want the red/green dots showing hexagonal symmetry to be bigger; 
  %when the figure is reduced they are hard to see.
  \caption{(a) R-spots increases in intensity when the temperature of the system is 
      lowered to 280K. (b) We measured the average angle made between each monomer 
      alkane tail and the membrane plane. The average tilt angle (dashed line) is 
      near -2\degree~which is far from the 37\degree~tilt angle previously used to
      explain R-spots. (c) To isolate the main cause of R-spots, we removed all atoms
      from the trajectory except for carbon atoms that constitute the tails. The
      simulated XRD pattern of the tails-only trajectory still shows R-spots. (d)
      Since the tails stay relatively flat, we plotted the centroid of each tail 
      and observed that they pack hexagonally (for example, green colored centroids
      in the plot surround the red centroid in hexagonal fashion). (e) We hypothesize
      that R-spots is the result of ordered tail packing. Defining
      the membrane plane to be 0\degree, we measured the angles between each alkane 
      chain tail centroid and its nearest neighbor centroids for the equilibrated 
      sandwiched configuration simulated at 280K. Peaks that appear in each distribution
      are centered near $\pm~33\degree$. (f) We radially integrated the simulated XRD
      patterns of the parallel displaced and sandwiched configuration within the region
      bounding R-alkanes. Peaks appear in the same location as the angle distributions
      which corroborates our hypothesis.}~\label{fig:tail_packing}
  \end{figure}

@@

  We can also produce R-double if layers are not uniformly spaced. Rather, monomers might form 
  pairs that stack less than 3.7~\AA~apart, and whose center of masses are spaced 7.4 \AA~ 
  from the next pair of monomers (Figure ~\ref{fig:staggered_rzplot_norestraints}). To our 
  knowledge, there have been no studies that specifically address the possibility of a
  configuration like this. 
  %MRS8: above: be more precise, for example, QM studies or crystallographic investigation of stacked pi-pi system, 
  % can cite the reference found.
  Our force field causes our system to tend towards uniformly spaced
  layers.   Simulations of unevenly spaced systems are only stable if position restraints
  are applied to heavy atoms of the phenyl rings.
  %MRS8: added below.
  There is little evidence from QM studies of stacked $\pi-\pi$ systems that this behavior is incorrect.~\ref{???}

  Adding water to the system 
  %facilitates 
  appears to facilitate 
  the appearance of R-double,
  %MRS8: added above. I realize the H-bonding isn't brought up until later paragraphs, but I don't want to leave people in suspense; it serves as essentially a thesis of the next few paragraphs, which do group together.
  through the creation of hydrogen bonding of water to carboxylate groups of consecutive layers.
  We added water to the 
  parallel displaced and sandwiched configurations in the ordered basin and equilibrated them
  according to the wet equilibration procedure. There is no experimental measurement of 
  water concentration in these membranes so we tested a range of water concentrations from
  1\% to 5\%. R-double appears transiently in the simulated XRD pattern of the parallel 
  displaced configuration with 1 wt\% water (Figure \ref{fig:solvated_pore_rzplot_norestraints}). It is
  not initially present, but appears after 200 ns of simulation time. After 450 ns, it 
  disappears again. Simulated XRD patterns of all other solvated systems tested are shown 
  in Figure S~\ref{S-fig:solvation}, however R-double is not present.

@@

  R-double appears in the solvated system due to the structure of the head groups. To 
  %prove 
  demonstrate
  this,
  we removed the head groups from the trajectory used to produce Figure~\ref{fig:solvated_pore_rzplot_norestraints}
  in order to produce that shown in Figure~\ref{fig:rdouble_nophenyls}. R-double does not appear
  without the influence of the head groups. Water molecules must play a role in the structuring
  of the head groups since R-double does not appear in any dry simulations.

@@

  %BJC6: moved this paragraph up to make the section more idea-focused
  In general, the composition of each region, particularly within our definition of 
  the pore region, is similar between all systems. 
  %MRS8: weakened a bit
  This suggest that the details of transport may be relatively independent of the structural differences between
  different possible structures studied here.
  %We believe we can solvate and study transport
  %in any of the systems presented here and extract similar transport mechanisms. We will 
  %need to verify that this assumption is true.

@@

  We calculated the ionic conductivity of the parallel displaced configuration in
  the ordered basin with 1 wt \% water since we believe its structure is the 
  closest match to experiment. Our model estimates the ionic conductivity within
  one order of magnitude of experiment (Figure~\ref{fig:ionic_conductivity}). We
  verified the value calculated by the Nernst-Einstein relationship using 
  the collective diffusion model\cite{liu_collective_2013} 
  (See section S~\ref{s-fig:conductivity}). The values calculated by the 
  %MRS8: this section reference not working either.
  collective diffusion model agree with Nernst-Einstein  values within error, 
  however there is a much higher uncertainty that would require much longer 
  simulations to lower.

@@

  % BJC5: Is it worth including the distance between layers and the correlation length? 
  % MRS8: probably very briefly discuss the changes, to justify the use of the word ``minor''
  There are only minor changes to the physical characteristics of these systems when 
  they are cross-linked (See Figure~\ref{fig:xlink}). The ionic conductivity of 
  the sandwiched configuration decreases while that of the parallel displaced 
  configuration stays the same. The pore spacing decreases in both systems by
  0.07 nm. The vertical monomer stacking distance increases in the sandwiched
  configuration and decreases in the parallel displaced configuration, however the 
  values of cross-linked configurations fall within uncertainty of their un-crosslinked
  counterparts. The correlation length decreases in both systems, 
  %MRS7: from what to what?
  %BJC6: If I say that, then shouldn't I say it for all metrics? That sounds
  % a bit clunky especially when they are plotted in the figure and I just want to make
  % somewhat general comments on the effect of cross-linking.
but is most pronounced in the parallel displaced system.

@@

  We achieve maximum structural consistency with experiment, as determined by simulated
  2D XRD patterns, when we build our model in the ordered basin parallel displaced
  configuration with 5 monomer columns per pore and 1 wt\% water. R-alkanes and R-pores
  appear where expected for the reasons originally predicted. We find that R-spots
  is likely due to ordered alkane chain packing. R-$\pi$ appears at lower q values than
  experiment because monomer stack too far apart. Finally, we observed that our model
  can only produce
  %reproduces 
  R-double when we add 1 wt\% water to the system. This observation has
  %important 
  possible
  implications for the reproducibility of the experimental results from Feng 
  et al since they claim that their membrane is completely dry.

@@

  %We justified that 
  Our system can reasonably estimate ionic conductivity.  Our
  calculations are about 1 order of magnitude higher than experiment, however
  that is to be expected since we are simulating a perfectly straight and
  defect-free membrane.

>>>>

  Nanostructured porous membranes made from the cross-linked hexagonal phase of
  self-assembled lyotropic liquid crystals (LLCs) are a promising material for
  selective separations. In this work, we investigate an experimentally
  characterized LLC membrane using an atomistic molecular model. We use
  simulated X-ray diffraction (XRD) patterns in order to ensure maximum
  consistency between our structures and experiment. We show that pores are
  likely composed of 5 columns of stacked LLC monomers which surround each
  hydrophilic core. Although this system has been reported as dry, we show that
  small amounts of water are necessary to fully reproduce the experimental XRD
  pattern due to asymmetries introduced by hydrogen bonds between monomer head
  groups and water molecules. We explore the composition and structure of the
  nanopores and reveal that there exists a composition gradient rather than a
  abrupt partition between the hydrophilic and hydrophobic regions. A caveat is
  that the time scales of the dynamics are extremely long for this system,
  requiring starting from a range of alternative initial configurations and
  careful examination of the metastable states observed. The clearer picture of
  the nanoscopic structure of these membranes provided in this study will enable
  a better understanding of the mechanisms of small molecule transport within
  these nanopores.
  \end{abstract}

@@

% MRS6: I would keep a paragraph about here how current membranes DON'T
% work for this purpose. But one paragraph is enough.
% BJC5: added back below paragraph.  
% MRS8: ``because they are unstructured'' lacks detail.  NF has some structure, since there are irregular pores. Check some other reviews and see the points they cover?
  Current commercial RO and NF membranes suffer limitations to their selectivity
  because they are unstructured. Although scalable, their fabrication involves the 
  spontaneous assembly of polymers into disordered structures which offer separation 
  pathways that are tortuous and polydisperse in size. This makes overcoming the well-known 
  permeability-selectivity tradeoff a challenge. Namely, it is difficult to increase
  the permeability of a desired molecular or atomic species, while maintaining the 
  same retention of an undesired species \cite{werber_materials_2016}.

@@

%MRS8: preliminary evidence has shown they may be?  It's more than a guess.
  Preliminary evidence has shown that crosslinked lyotropic liquid crystal
  (LLC) membranes may be capable of performing highly selective separations. LLCs
  are amphiphilic molecules that have the ability to self-assemble into porous
  nanostructures \cite{smith_ordered_1997} and can be crosslinked to create
  mechanically strong membrane films with pores on the order of 1 nm in diameter
  \cite{zhou_supported_2005}. Unlike most commercial NF membranes, LLC membrane
  pores are uniform in size because they are self-assembled.  Since LLC membranes
  lack a pore size distribution, they inherently exhibit high selectivity due to
  their strict molecular weight cut-off (MWCO)~\cite{zhou_supported_2005}.
  Additionally, the LLC monomers examined in this paper are salts, and therefore
  lead to Donnan exclusion of ions in solution. The membrane gains a net surface
  charge when counterions from the head groups that line the pore walls escape
  into the feed solution in an effort to balance the gradients of concentration
  and electric potential \cite{donnan_theory_1995}.

@@

  Recently, researchers have learned how to macroscopically align the hexagonal
  domains which has revived research into H\textsubscript{II} phase LLC membranes.
  Previously, research efforts were focused on the Q\textsubscript{I} phase,
  whose geometry does not require alignment. In 2014, Feng et al.~showed that one
  can align Col\textsubscript{h} hexagonal domains, created by the monomer Na-GA3C11,
  using a magnetic field with subsequent cross-linking to lock the structure in 
  place\cite{feng_scalable_2014}. In 2016, Feng et al. showed that one could obtain
  the same result by confining neat monomer between PDMS or glass substrates since 
  hexagonal mesophases preferentially anchor perpendicular to both surfaces\cite{feng_thin_2016}.
%  the same result using a second technique termed soft confinement\cite{feng_thin_2016}.
  %MRS8: maybe say what it is, not just what it's named? 
  Current experimental efforts are focused on extending the method to the 
  H\textsubscript{II} phase and characterizing the performance of these newly aligned systems.

@@

  In this study, we achieve a  more realistic atomistic description of LLC
  membranes than, to our knowledge, has ever previously been created, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the results using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure drawn from small angle X-ray scattering (SAXS) and wide angle
  X-ray scattering (WAXS) experiments as well as in matching ionic conductivity
  measurements \cite{feng_thin_2016}.

@@

  In this paper, we perform molecular modeling of the Col\textsubscript{h}
  assembly formed by Na-GA3C11. Compared to the H\textsubscript{II} phase, the
  Col\textsubscript{h} phase is a simpler starting point. The system is not
  assembled in aqueous phase which allowed us to simulate longer timescales, and
  there exists detailed experimental characterization of the fully aligned state,
  including 2D wide-angle X-ray scattering (WAXS) patterns
  (Figure~\ref{fig:WAXS}) which are useful for reconstructing structural data.

  %MRS8: I started tweaking the below, but it might be better just to eliminate.
  % Modeling of Col\textsubscript{h} will help in modeling 
  %H\textsubscript{II} in the future.
  %The two phases appear to share similar structural characteristics since the
  %pore spacings in each system are in close agreement
  %\cite{feng_thin_2016,resel_structural_2000}.

@@

 Despite having structural data, there is still information which experiment
 cannot definitively answer. Specifically, we want to know:
 \begin{enumerate}
    \item What is the density of monomers that pack around each hydrophilic core? 
    \label{point:monomernum}

@@

  	We are careful to avoid the term 'layers' since many liquid crystalline systems
 	have long range order in 1 or 2 spatial dimensions and short range order in 	
 	the other dimensions.~\cite{chaikin_principles_1995}  % page 58
 	In the system we are studying, there are long-range 2D correlations in the 
 	hexagonal array of pores (xy plane) and short range z-direction correlations between 
        stacked columns of monomers. In this study, we use atomistic molecular modeling
        to study how the system's structure is affected by the density of these monomer columns
        that surround each pore's hydrophilic core.

@@

	While the Col\textsubscript{h} phase is described as dry, it is likely
	that small amounts of ambient water are absorbed into the system. The hydrogen
	bonding network formed by the water may play a role in structuring the pore. We
	used simulated X-ray diffraction (XRD) patterns to see if there is any
	meaningful structural difference between a "dry" and a "wet" system.

	\item What is the detailed atomistic structure of the pores?\label{point:composition}

@@

  \subsection{Equilibration}

  We developed equilibration schemes for creating dry and wet configurations.
  Both schemes start with an initial configuration generated according to the
  previous guidelines. For wet systems, we added the desired concentration of
  water to the initial configuration and carried out equilibration in the same
  way as the dry systems. First, we fixed monomer head groups in place using
  position restraints with a force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We
  gradually released the position restraints by decreasing the force constants
  over a series of NVT simulations. We allowed the resulting unrestrained
  structure to equilibrate for 5 ns in the NPT ensemble with pressure controlled
  by the berendsen barostat followed by NPT equilibration simulations run for at
  least 400 ns using the Parrinello-Rahman barostat. More equilibration details
  are available in section S-\ref{S-section:equilibration}.

@@

  Our simulations best support a model built with 5 monomer columns per pore based on
  the measured equilibrated pore-to-pore distances. To discern the composition of
  the monomer layers, addressing question \ref{point:monomernum} above, we ran
  simulations of systems created with 4--8 columns per pore. We built systems in both
  the parallel displaced and sandwiched configurations and equilibrated them according
  to the dry equilibration procedure. We tested all systems with an initial vertical monomer 
  spacing, $\mathit{d}$, of 3.7 \AA~in accordance with R-$\pi$. We tested 4 additional 
  systems with monomers initially spaced 5 \AA~apart vertically (see section~\ref{S-section:initial_pore_spacing}
  of the Supporting Information for more details on sensitivity to initial layer spacing). 
  We considered the pore-to-pore spacing to be equilibrated as defined in section~\ref{method:equil_time}.
  Figure~\ref{fig:p2p} shows the equilibrated pore-to-pore distances for all systems tested.

@@

  All systems tested, although equilibrated from the perspective of the metrics used here, are
  frozen in metastable basins. Not all make physical sense or fit the experimental profile that 
  we are trying to match. In the limit of infinite simulation time, all systems will in theory 
  converge to a single equilibrium configuration, but that time is far beyond the 100's of 
  nanosecond simulated here. For simplicity, we group the systems studied here into the ordered
  and disordered basins. What we find is that any system where $\mathit{d}=3.7$ \AA~can generally 
  be characterized as being in an
  more ordered basin, and any system where $\mathit{d}=5.0$ \AA~can be characterized as a more the disordered basin.
  Generally, when monomers are started further apart, they stay further apart than systems where
  monomers are started closer together (see Table~\ref{table:correlation_length}). Because of the
  extra space between stacked monomers, monomer head groups have more rotational freedom. We
  quantified the ordering of the head groups using the nematic order parameter (see section S-\ref{S-method:nematic_order} for details of calculation). 
  Disordered basin systems have a lower nematic order parameter (meaning they are more disordered) 
  than ordered basin systems.

@@

  To understand the most likely origin of R-spots, we determined which atoms
  gave rise to the feature. 
  %MRS8: next two sentences repetitive; could you revise? 
%  Since R-spots is present as higher intensity spots within R-alkanes, it is
%  likely that the spots arise as a consequence of the tails. 
  By removing all non-tail atoms from the trajectory and simulating a
  diffraction pattern with the remaining atoms, we were able to isolate the cause
  of the spots to the tails (Figure~\ref{fig:tails_rzplot}). Since the tails stay
  nearly flat, we plotted their centroids (Figure \ref{fig:centroids}) and
  measured the angle between each centroid and its nearest neighbors with respect
  to the plane of the membrane. We see distinct peaks in the distribution of
  these angles (Figure~\ref{fig:tail_packing}).

@@

  \begin{figure}[!htb]
  \begin{subfigure}{0.32\linewidth}
  	\centering
  	\includegraphics[width=\textwidth]{rzplot_layered_280K_jet.png}
  	\caption{}\label{fig:sandwiched280K}
  \end{subfigure}
  \begin{subfigure}{0.32\linewidth}
    \includegraphics[width=\textwidth]{tilt_dist.png}
    \caption{}\label{fig:tilt}
  \end{subfigure}
  \begin{subfigure}{0.32\linewidth}
	\centering
	\includegraphics[width=\textwidth]{tails_rzplot_jet.png}
	\caption{}\label{fig:tails_rzplot}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\linewidth}
    \centering
	\includegraphics[width=\textwidth]{centroids_box.png}
  \caption{}\label{fig:centroids}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\linewidth}
        \centering
	        \includegraphics[width=\linewidth]{angles_traj_layered.png}
	        \caption{}~\label{fig:layered_tails}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
        	\centering
	        \includegraphics[width=\linewidth]{layered_angle_v_I.png}
	        \caption{}~\label{fig:layered_integration}
  \end{subfigure}
  \caption{(a) R-spots increases in intensity when the temperature of the system is 
      lowered to 280K. (b) We measured the average angle made between each monomer 
      alkane tail and the membrane plane. The average tilt angle (dashed line) is 
      near -2\degree~which is far from the 37\degree~tilt angle previously used to
      explain R-spots. (c) To isolate the main cause of R-spots, we removed all atoms
      from the trajectory except for carbon atoms that constitute the tails. The
      simulated XRD pattern of the tails-only trajectory still shows R-spots. (d)
      Since the tails stay relatively flat, we plotted the centroid of each tail 
      and observed that they pack hexagonally (for example, green colored centroids
      in the plot surround the red centroid in hexagonal fashion). (e) We hypothesize
      that R-spots is the result of ordered tail packing. Defining
      the membrane plane to be 0\degree, we measured the angles between each alkane 
      chain tail centroid and its nearest neighbor centroids for the equilibrated 
      sandwiched configuration simulated at 280K. Peaks that appear in each distribution
      are centered near $\pm~33\degree$. (f) We radially integrated the simulated XRD
      patterns of the parallel displaced and sandwiched configuration within the region
      bounding R-alkanes. Peaks appear in the same location as the angle distributions
      which corroborates our hypothesis.}~\label{fig:tail_packing}
  \end{figure}

@@

  We can also produce R-double if layers are not uniformly spaced. Rather,
  monomers might form pairs that stack less than 3.7~\AA~apart, and whose center
  of masses are spaced 7.4 \AA~ from the next pair of monomers (Figure
  ~\ref{fig:staggered_rzplot_norestraints}). To our knowledge, there have been no
  studies that specifically address the possibility of a configuration like this.
  Our force field causes our system to tend towards uniformly spaced layers.
  Simulations of unevenly spaced systems are only stable if position restraints
  are applied to heavy atoms of the phenyl rings.  There is little evidence from
  QM studies of stacked $\pi-\pi$ systems that this behavior is
  incorrect.~\ref{tauer_beyond_2005}

  Adding water to the system appears to facilitate the appearance of R-double,
  through the creation of hydrogen bonding of water to carboxylate groups of
  consecutive layers. We added water to the parallel displaced and sandwiched
  configurations in the ordered basin and equilibrated them according to the wet
  equilibration procedure. There is no experimental measurement of water
  concentration in these membranes so we tested a range of water concentrations
  from 1\% to 5\%. R-double appears transiently in the simulated XRD pattern of
  the parallel displaced configuration with 1 wt\% water (Figure
  \ref{fig:solvated_pore_rzplot_norestraints}). It is not initially present, but
  appears after 200 ns of simulation time. After 450 ns, it disappears again.
  Simulated XRD patterns of all other solvated systems tested are shown in Figure
  S~\ref{S-fig:solvation}, however R-double is not present.

@@

  R-double appears in the solvated system due to the structure of the head
  groups. To demonstrate this, we removed the head groups from the trajectory
  used to produce Figure~\ref{fig:solvated_pore_rzplot_norestraints} in order to
  produce that shown in Figure~\ref{fig:rdouble_nophenyls}. R-double does not
  appear without the influence of the head groups. Water molecules must play a
  role in the structuring of the head groups since R-double does not appear in
  any dry simulations.

@@

  %BJC6: moved this paragraph up to make the section more idea-focused
  In general, the composition of each region, particularly within our
  definition of the pore region, is similar between all systems.  This suggests
  that the details of transport may be relatively independent of the structural
  differences between different possible structures studied here.
  %We believe we can solvate and study transport
  %in any of the systems presented here and extract similar transport mechanisms. We will 
  %need to verify that this assumption is true.

@@

  We calculated the ionic conductivity of the parallel displaced configuration in
  the ordered basin with 1 wt \% water since we believe its structure is the 
  closest match to experiment. Our model estimates the ionic conductivity within
  one order of magnitude of experiment (Figure~\ref{fig:ionic_conductivity}). We
  verified the value calculated by the Nernst-Einstein relationship using 
  the collective diffusion model\cite{liu_collective_2013} 
  (See section S~\ref{S-fig:conductivity}). The values calculated by the 
  collective diffusion model agree with Nernst-Einstein  values within error, 
  however there is a much higher uncertainty that would require much longer 
  simulations to lower.

@@

  % BJC5: Is it worth including the distance between layers and the correlation length? 
  % MRS8: probably very briefly discuss the changes, to justify the use of the word ``minor''
  There are only minor changes to the physical characteristics of these systems when 
  they are cross-linked (See Figure~\ref{fig:xlink}). The ionic conductivity of 
  the sandwiched configuration decreases while that of the parallel displaced 
  configuration stays the same. The pore spacing decreases in both systems by
  0.07 nm. The vertical monomer stacking distance increases in the sandwiched
  configuration and decreases in the parallel displaced configuration, however the 
  values of cross-linked configurations fall within uncertainty of their un-crosslinked
  counterparts. The correlation length decreases in both systems, 
  but is most pronounced in the parallel displaced system.

@@

  We achieve maximum structural consistency with experiment, as determined by simulated
  2D XRD patterns, when we build our model in the ordered basin parallel displaced
  configuration with 5 monomer columns per pore and 1 wt\% water. R-alkanes and R-pores
  appear where expected for the reasons originally predicted. We find that R-spots
  is likely due to ordered alkane chain packing. R-$\pi$ appears at lower q values than
  experiment because monomer stack too far apart. Finally, we observed that our model
  can only produce R-double when we add 1 wt\% water to the system. This observation has
  possible implications for the reproducibility of the experimental results from Feng 
  et al. since they claim that their membrane is completely dry.

@@

  Our system can reasonably estimate ionic conductivity. Our
  calculations are about 1 order of magnitude higher than experiment, however
  that is to be expected since we are simulating a perfectly straight and
  defect-free membrane.

  Finally, we verified that our conclusions do not change when the system is
  cross-linked by the algorithm we implemented. The diffraction pattern weakens
  relative to the uncross-linked system, the ionic conductivity drops by a factor
  of ca. 1.5, in closer agreement with experiment, the pore spacing decreases and
  the membrane becomes thicker; however, all changes are relatively minor, 
  preserving most of the features well.
